### 2024-06-15 getting-started.qmd
20240608 - packages to install
20240614 - wide vs long form

install.packages(c("petersenlab","tidyverse","psych"))
```

@@

Here are the data in the `nfl_actualStats_offense_weekly` object.
The data are structured in player-season-week form.
That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.

@@

Below, we widen the data widened by two variables (`season` and `week`), using `tidyverse`, so that the data are now in player form (where each row is uniquely identified by the `player_id` variable):

@@

>>>>

install.packages(
  c("petersenlab","ffanalytics","nflreadr","nflfastR","nfl4th","nflplotR",
  "gsisdecoder","progressr","lubridate","tidyverse","psych"))
```

@@

Depending on the analysis, it may be important to restructure the data to be in long or wide form.
When the data are in wide form, each player has only one row.
When the data are in long form, each player has multiple rows—e.g., a row for each game.
The data structure is called wide or long form because a dataset in wide form has more columns and fewer rows (i.e., it appears wider and shorter), whereas a dataset in long form has more rows and fewer columns (i.e., it appears narrower and taller).

Here are the data in the `nfl_actualStats_offense_weekly` object.
The data are structured in "player-season-week form".
That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.
This is an example of long form, because each player has multiple rows.

@@

Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "player form" (where each row is uniquely identified by the `player_id` variable):

@@

Conversely, we can also restructure data from wide to long.
